---
title: "Lab 6"
author: "Corinna Hong"
date: "November 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in penguins.csv data
```{r, include = FALSE}

library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")

```

Is there a significant difference in mean flipper length for female penguins by species? (Chinstrap, Gentoo, Adelie)

2. Do some exploring

New functions: with + table for easy counting over multiple factors

Update species name using case_when

```{r}

counts <- with(penguins, table(species,sex)) # with is telling R, "with this data, do this" with penguins, make a table of counts for species and sex 

penguins2 <- penguins %>%
  mutate(common_name =
           case_when(
             species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
             species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
             species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
           )) %>% 
  select(-species) %>% # Take out the column species
  filter(sex == "FEMALE") # filter female data from sex column
  
```

- Exploratory graphs
- Summary statistics
- Levene's test for equal vcarinces
- One-way ANOVA
- Post-hoc testing

```{r}

hists <- ggplot(penguins2, aes(x= flipper_length)) +
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~common_name)
hists


qq <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq()+
  facet_wrap(~common_name)
qq


# Levene's test for equal variances (>2 groups). If you have two gorups, use the F-test for equal variances
# We'll use the function leveneTest from the 'car' package
# H0: Variances are equal
# HA: Variances are NOT equal

penguin_levene <- leveneTest(flipper_length ~ common_name, data = penguins2)
penguin_levene

# We retain the null hypothesis of equal variances (p = 0.06)


# What are the actual variances?

variance_table <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )
variance_table
```

3. One-way ANOVA

Single factor: common_name (species)
Number of levels in the factor: 3 (Chinstrap, Gentoo, Adelie)
Random variable: flipper_length (mm)

Is there a significant diference in mean flipper length (mm) for female Chinstrap, Gentoo, and Adelie penguins?

```{r}

# HO: Mean flipper length across all species are equal
# HA: AT LEAST two means differ significantly

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2) # COmpare means of this random variable, grouped by factor, from this data
summary(penguin_aov) # Use summary() to get it to tell you more information

# P-Value: If we take samples from these three gorups and they come form populations with the same mean fliper lenght, there is almost a 0% chance that we are going to get sample means at least this different by random chance

# At least two samples were taken from populations with different means. Which ones are different?


# Post-hoc using Tukey's HSD

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

```

Mean flipper lengths for Adelie, Chinstrap, Gentoo penguins are significantly different (F(df) = statistic, p < 0.001, $\alpha$ = 0.05) with post-hoc Tukey's HSD, $\alpha = 0.05$

